---
title: "A simple PEC chatbot"
format: html
embed-resources: true
---

```{r}
#| label: setup
#| message: false

library(ellmer)
library(tidyverse)
```


```{r}
#| label: the r counter function

r_counter <- function(word){
  letters <- str_split_1(word, "")
  sum(letters == "r") + sum(letters == "R")
}
```

A generic call to an LLM won't suffice for many tasks.

```{r}
#| label: generic response

generic <- chat_anthropic()

story1 <- generic$chat("Write a 350 (approx) word romance that includes the letter r exactly 50 times.",
                       echo = FALSE)
story1

r_counter(story1) # usually wrong
```

## The planner-executor-critic model

By assigning specific tasks to multiple different chats, we can often do much better. Here we'll create 4: a user-facing control bot, a planner, and executor, and a critic. 

```{r}
#| label: chat creation

chatbot <- chat_anthropic(system_prompt = "You are a user-facing chatbot which provides high quality solutions to complex tasks. Use the provided tools to follow a planner-executor-critic model. If the critic returns an evaluation below 9/10, rerun the cycle to reflect its feedback. In this situation, pass along the critic's revisions to the executor (or, if absolutely necessary, to the planner) EXACTLY, adding only an instruction that the executor MODIFY its previous attempt per these revision. It should NOT start over from scratch. Do NOT add additional edits of your own. Iterate this process a maximum of THREE times, total, and return the best execution. NEVER directly revise or even evaluate the executor's responses.")

planner <- chat_anthropic(system_prompt = "You are a planner. Your job is to turn the user's prompt into a specific, ordered task list which another LLM will later use to complete the request in a VERY high quality manner. Do NOT execute the task list or attempt to fulfill the prompt. Instead, return a prompt for an executor LLM.")

executor <- chat_anthropic(system_prompt = "You are an executor. Your job is to follow the steps provided EXACTLY, returning a COMPLETE and ACCURATE response to the request.")

critic <- chat_anthropic(system_prompt = "You are an evaluator. Your job is to compare the prompt provided and the response generated and decide if the task has been completed satisfactorily. Provide an integer evaluation from 1 (worst) to 10 (best). For scores lower than 9, provide a SPECIFIC plan for MODIFYING the response to meet the demands of the prompt. Do NOT provide instructions for starting over.")
```

Next we create and register tools so that the user-facing chatbot can make use of them.

```{r}
#| label: tool creation

create_plan <- function(prompt){
  planner$chat(prompt, 
               echo = FALSE)
}

execute_plan <- function(plan){
  executor$chat(plan, 
                echo = FALSE)
}

evaluate_response <- function(prompt, response){
    critic$chat(paste("Prompt:", prompt, "Response:", response),
              echo = FALSE)
}

create_plan_tool <- 
  tool(create_plan, 
       description = "use the planner LLM to create a specific plan for fulfilling the user's request.",
       arguments = list(prompt = type_string("the user's request.")))

execute_plan_tool <- 
  tool(execute_plan, 
       description = "use the executor LLM to carry out the execution plan for the user's prompt.",
       arguments = list(plan = type_string("the plan to be carried out.")))

evaluate_response_tool <- 
  tool(evaluate_response, 
       description = "use the critic LLM to evaluate the executor's fulfillment of the plan.",
       arguments = list(prompt = type_string("the user's original prompt."),
                        response = type_string("the response to be evaluated.")))

tools <- c(create_plan_tool,
           execute_plan_tool,
           evaluate_response_tool)

chatbot$register_tools(tools)
```

## Testing it out

In this example, we ask for a very specific count of one letter. Let's provide a tool to the critic for checking that this requirement is satisfied by the executor.

```{r}
#| label: r counter tool

r_counter_tool <- 
  tool(r_counter, 
       description = "count the number of r's in a string",
       arguments = list(word = type_string("the string to consider")))

critic$register_tool(r_counter_tool)
```

Let's see this in action.

```{r}
#| label: final product

chatbot$chat("Write a 350 (approx) word romance that includes the letter r exactly 50 times.")

chatbot$get_cost()
planner$get_cost()
executor$get_cost()
critic$get_cost()
```

Highly inefficient, sure, but not hard to improve. 